---
title: "Untitled"
format: html
editor: visual
---

# Question

A study is being designed to test the effectiveness of a new weight loss drug.  Subjects will be weighed before the drug is administered and again three weeks afterward.  The researcher wants the study to be sensitive enough to detect a weight loss of about 1 kg, and needs to know if the study is viable with 25 subjects. The only paper the investigators could find describing a similar study using a different weight loss program but also over three weeks writes “The 14 subjects lost an average of 1.8 kg (p=0.010)”. 

Using only this information, this quiz walks through a draft of a message to the investigator with an opinion on whether the study is viable, providing an appropriate power estimate in the process. 

## Email

Hello,

Assuming a similar two-sided one-sample (paired) t-test was performed in the other study, we can determine that the standard deviation in that study was about  [1.029]{.underline}. Let me know if you’d like the details on how I arrived at this figure.

Assuming that our population has the same standard deviation, the same design, no dropouts, and a 5% significance level, your proposed study with 25 participants has     B   % power to detect a change of 1 kg. Therefore, your proposed study would be    C     to detect your anticipated effect of 1 kg.

With 25 subjects, you would have 80% power to detect a difference of      D     kg. If you need to stick to a 1 kg detectible effect, then you’d want to recruit at least    E      subjects in order to have 80% power.

A. 2.24

B. 57.23

C. Underpowered

D. 1.31

E. 42

```{r}
library(pwr)
```

```{r}

# My code

pwr.t.test(n = 25, d = 0.4464, sig.level = 0.05, power = NULL, type = "paired")

# Parameters
n <- 25          # Number of subjects
sd <- 2.24       # Standard deviation from previous study
alpha <- 0.05    # Significance level
power <- 0.80    # Desired power

# Calculate the effect size
effect_size <- pwr.t.test(n = NULL, d = 0.4464, sig.level = alpha, power= .80, type = "paired")$d
effect_size

# Convert effect size to detectable difference
detectable_difference <- effect_size * sd

detectable_difference

```

# Do Over

1\)

```{r}
library(pwr)
```

```{r}
effect_size <- pwr.t.test(n = 25, d = 1, sig.level = 0.05, type = "paired")
effect_size



sd <- 2.24
# Parameters
n <- 25          # Number of subjects
sd <- 0.922      # Standard deviation from previous study
alpha <- 0.05    # Significance level
power <- 0.80    # Desired power

```
